RadLib: a radiative property model library for combustion CFD

### 2021-07-24 submissions/CPC_response_to_reviewers.md

	**The CMake build system has been updated such that ```ARCHIVE DESTINATION``` is explicitly specified, so that error should no longer occur. A build option has been added that makes the Python interface installation optional. Note that the C++ source code is required to install the Python interface, so there is no option to build only the Python interface. All of this information has been added to the ```README.md``` files located in the root directory and the ```build``` directory along with a more thorough list of software dependencies and required Python packages. Furthermore, we have now done full tests of the installation and running of all examples on a desktop Mac, a desktop Linux machine, and a large Linux-based cluster.**

2) The library includes PM, one WSGG, and a RCSLW model. The authors should clearly indicate the rationale of these choices (i.e., why only these three models) early on in the manuscript. Since this is intended for researchers who, presumably, are not experts in radiation modeling, the inclusion of these three models can be misleading. One might think that these three models are the most appropriate models for their simulations.

	**Section 2 has been heavily revised and should better address this issue. In particular, at the beginning of section 2, we note the why these models were chosen. The three models are practically relevant, and represent a variation in complexity and accuracy.**

@@

3) RadLib is intended to be expanded with new radiation property models in the future and interfaced with a user-supplied solver module for RTE. A brief discussion on how to interface such an external solver module should be presented. There is an example of a ray-tracing solver presented in the code files, but a brief discussion on coupling one's own RTE solver is important in the manuscript. Additionally, a brief discussion on how a user can add a new radiative property model should also be included.

	**Section 3 has been revised and extended to include more information on how users can include and interact with the RadLib library, including explicit references to example files that further illustrate usage of the library. We have also added a subsection to the results discussion that describes the integration of RadLib with the Fire Dynamics Simulator (FDS) code developed by NIST. Adding additional radiative property models to RadLib would require creating another class object for the implementation. The intention would be for the new model to inherit from the ```rad``` base class and reimplement the interface functions ```get_k_a``` and ```get_k_a_oneband```. Depending on the model, additional functionality may be needed. The Fortran and Python interfaces are wrapper functions that can be extended by paralleling the existing code, which is much simpler than writing these interfaces from scratch. A discussion on these points has been added to Section 3.**

4) The Planck-mean absorption coefficient included in RadLib is based on TNF workshop data (from 2003). According to the references cited, these are fairly old correlations. If the authors use these data, I am concerned that the model parameters and correlations used in RadLib for the PM model may be obsolete and inaccurate. Can the authors please comment on this?

	**We purposely chose the TNF workshop data for calculating Planck Mean absorption coefficients because it is one of the most common and widely used radiation models applied to turbulent combustion simulations, despite its age and limitations. Section 2.1 has been revised to include a more explicit dicsussion of why the TNF radiation model was chosen. Sections 4 and 5 include example cases and results that demonstrate the deficiencies of the PM model compared to the other models included, and some additional discussion is given there. We intend to extend the model to include updated PM data.**

### 2021-07-24 manuscript/manuscript.tex
DOL proofread and edit the paper. Rewrote the PM section. Checked for consistency with PM issues with reviewer 3

%% This list environment is used for the references in the
%% Program Summary
%%
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{float}
\usepackage[version=4]{mhchem}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{seqsplit}

@@

\newcommand{\BS}{\boldsymbol}

@@

\begin{abstract}
RadLib is a C++ library of radiation property models that can be applied to variety of systems involving radiative heat transfer, including CFD simulations. RadLib includes three major radiation property models---Planck Mean (PM) absorption coefficients, the weighted sum of gray gases (WSGG) model, and the rank-correlation spectral line weighted-sum-of-gray-gases (RCSLW) model---but its design and C++ and Python interface options permit convenient expansion to additional models. Several example cases illustrate the use of the models with an included ray-tracing solver and compare them in terms of accuracy relative to line-by-line solutions. The computational cost of the models is also compared. RadLib provides researchers with convenient access to validated radiation property models and a framework for further development. 
\end{abstract}

@@

% PROGRAM SUMMARY.

{\bf PROGRAM SUMMARY}
  %Delete as appropriate.

@@

Combustion CFD is uniquely challenging because it involves numerous tightly coupled physical phenomena---including multicomponent mass transfer, convective and radiative heat transfer, turbulent fluid dynamics, multi-phase and particle flows, and complex chemical reaction kinetics---that span many orders of magnitude in both length and time scales. Specifically, combustion simulations are often complicated by a lack of accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost. Historically, radiative heat transfer modeling in combustion problems is either neglected or oversimplified, resulting in significant over- or underprediction of flame temperatures by up to 200$^{\circ}$C, respectively \citep{Modest_2016}. Additionally, some late-stage flame phenomena, particularly soot behavior and flame sheet breakthrough, increasingly depend on and influence the magnitude of radiative heat losses and are difficult to simulate accurately without robust radiation modeling \cite{Modest_2013}. These problems motivated development of the RadLib library, which implements three radiation property models in C++ and Python and provides a framework for additional models and further development, giving researchers more convenient access to advanced radiation property modeling tools. 
%
%Modeling radiative heat transfer for CFD is complex and challenging. For example, combustion simulations are often complicated by the need for accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost. This can lead to simulations that use oversimplified radiation models, but in cases where radiation is significant relative to the other physical phenomena, oversimplified models can negatively impact the accuracy of simulation results. In some cases, radiative gains or losses are small compared to other energy sources or heat transfer modes and radiation can be safely neglected, but many practical engineering systems do not allow this. For example, simple jet flames demonstrate both possibilities: early in a flame's development, turbulent mixing typically dominates heat transfer and dictates flame behavior; late-stage flame phenomena like soot behavior and flame sheet breakthrough, however, increasingly depend on the magnitude of radiative heat losses and are difficult to simulate accurately without a robust radiation model.

The core mechanisms of radiative phenomena are physically and mathematically complex. Unlike other modes of heat transfer, radiation is governed by a set of integro-differential equations that depend on both direction and wavelength, known as the radiative transfer equations (RTE). Computational radiation solvers require two distinct but interconnected parts: (1) a solution approach for the radiative transfer equations (RTE) and (2) a radiation property model. 
Complexity, implementation, and computational costs of these two parts can vary widely, depending on the models chosen. The optically thin model for solution of the RTE is trivial, while the discrete ordinates model (DOM) and Reverse Monte Carlo Ray Tracing (RMCRC) model, among others, are significantly more involved. Similarly for radiation property models, which range from simple correlations for mean absorption coefficients, to gray gas models such as the weighted sum of gray gas model (WSGG), to the more advanced Spectral Line WSGG (SLW) family of model, to full line-by-line (LBL) approaches. In practical engineering simulations, the solution of the RTE is often the most complex and costly part of radiative simulation, but advanced property models, such as the Rank Correlated SLW (RCSLW), discussed here, are nontrivial models to understand, implement, and validate.
RadLib addresses this second, radiative property model part of radiative simulation by providing a set of implemented and validated radiation property models that can be applied to various solution approaches for the RTE.

For an absorbing, emitting, and non-scattering medium, the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$. Solving the RTE requires information about the radiative properties of the materials and media in the system. Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model. Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules, reflecting varying degrees of dependence on local system properties like temperature, pressure, and gas composition. Radiation property models, including those implemented in RadLib, typically use high-resolution spectral databases like HITRAN and HITEMP \cite{Rothman_2010} to calculate spectral absorption coefficients, reducing or simplifying the huge amount of spectral data using correlations and modeling assumptions \citep{Zhang_2002b}.

Direct solution of the spectral RTE involves so-called line-by-line calculations in which the RTE is solved for each individual spectral line, potentially millions of evaluations; this is computationally prohibitive for most practical configurations and used primarily for model validation. Traditional band models rely on estimated or averaged "line-of-sight" property values that, as a result, cannot be incorporated into the RTE. Modern band models require high-resolution databases and reorder rather than averaging absorption coefficients, but tend to lose accuracy when applied to nonhomogeneous media and may still require large numbers of RTE evaluations \citep{Modest_2016}.

Global models are an important class of radiation property models that use spectrally-integrated radiation properties \cite{Hottel_1967,Modest_2013,Modest_2016}. Assuming spectrally uniform (gray) properties over a region $j$ (that is, for a gray gas $j$), we can integrate Eq.~\ref{e:RTE_general} over the fixed wavenumber ranges $\eta_j$ defined by $j$ to obtain 
%
\begin{linenomath}
\begin{equation} \label{e:RTE_wsgg}
	\frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
\end{linenomath}
%
where $I_b$ is the blackbody intensity, $I_j$ is the radiative intensity in the region defined by gray gas $j$, $\kappa_j$ is the spectrally integrated absorption coefficient for gray gas $j$, and $a_j$ is a weighting factor defined by
%
\begin{linenomath}
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
\end{linenomath}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~\ref{e:RTE_wsgg} can be evaluated for each of the $n$ gray gases with the addition of one "clear gas" $j=0$ that represents transparent windows in the spectrum for which $\kappa_0=0$. This "gray gas" approach allows us to calculate the total radiative intensity $I$ by summing the individual gray gas intensities rather than integrating over the entire wavenumber spectrum: 
%
\begin{linenomath}
\begin{equation}
	I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
\end{linenomath}
%
%The heat flux $\mathbf{q}$ and volumetric radiative heat source $Q$ can then be calculated by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere and finding the divergence or the resulting vector field, respectively:
%%
%\begin{linenomath}
%\begin{equation}
%    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
%\end{equation}
%\end{linenomath}
%%
%\begin{linenomath}
%\begin{equation}
%    Q = -\nabla\cdot\mathbf{q}.
%\end{equation}
%\end{linenomath}
%%
Global models are especially attractive to computational combustion researchers because they significantly reduce the required number of RTE evaluations during calculations of radiative source terms for the energy conservation equations.

With RadLib, we aim to provide researchers with a convenient access point for radiative property models that can be interfaced with various RTE solution approaches suitable for CFD applications. RadLib is a C++ library with an optional Python interface. Currently, RadLib only calculates radiation absorption coefficients and weighting factors for use in global radiation models, but expansion to other radiative properties and model types is possible. Three models are fully implemented and validated. All models use a common interface to facilitate researchers' access to and use of radiation modeling tools suitable for combustion CFD applications.

@@

At present, RadLib includes three radiation property models of varying complexity: Planck Mean absorption coefficients using the optically thin approximation, a weighted sum of gray gases (WSGG) model, and the rank-correlated spectral line weighted-sum-of-gray-gases (RCSLW) model. Each of these models was specifically chosen for its notable attributes such as popularity, ease of implementation, or accuracy; their individual strengths, weaknesses, and limitations are discussed in the sections that follow.

Absorption coefficients and other radiative property data may take on different forms depending on the RTE solution approach, but can generally be classified as either spectral (wavenumber-dependent) or global (integrated over wavenumber). Spectral absorption coefficients can be applied to the spectral RTE (Eq.~\ref{e:RTE_general}), while global absorption coefficients are used in the spectrally-integrated RTE (Eq.~\ref{e:RTE_wsgg}). The type of absorption coefficient generated by each model and potential RTE integration strategies will also be discussed in the following sections. While solving the RTE is not the focus of this software, RadLib's example cases do employ a simple ray tracing solver to calculate the radiative heat flux and volumetric heat source profiles between two parallel planes. Refer to the example cases in Section \ref{s:Examples} for details.

Radiation absorption coefficients are typically calculated via correlations developed from high-resolution spectral databases like RADCAL, HITRAN, and HITEMP \citep{Grosshandler_1993,Rothman_2010} that relate spectral line strength to properties such as temperature, pressure, and gas composition \citep{Zhang_2002b}. At present, RadLib considers up to four gas species (H$_2$O, CO, CO$_2$, and CH$_4$), depending on the model, and soot. Note that the property models implemented in RadLib apply only to non-scattering media.

@@

RadLib's simplest radiation property model generates Planck Mean absorption coefficients under the assumption of optically thin media, from here on referred to as the Planck Mean (PM) model.
The optically thin approximation assumes that there is little to no reabsorption of radiative energy in the system, or, in other words, that the system only emits radiation and does not absorb it. In combustion systems, optical thinness is associated with various other flame properties such as low radiative heat transfer relative to chemical heat release or convective heat transfer, negligible amounts of soot or other large molecules, and relatively small flame lengths. These flame attributes are not universally indicative of optical thinness, but are often used to justify neglecting or simplifying radiative heat transfer modeling. Assuming optical thinness, however, has the distinct advantage of simplifying the radiative source term (denoted $S_{rad}$) of the energy conservation equation to 
\begin{equation}
	S_{rad}=-4\kappa \sigma T^4,
\end{equation}
where $\kappa$ is the radiation absorption coefficient, $\sigma$ is the Stefan-Boltzmann constant, and $T$ is temperature. Using an optically thin approximation means that no RTE solution is required when calculating radiative source terms.

In combustion systems, this simplification is convenient because it makes radiation modeling easy to implement and computationally inexpensive. The vast majority of computational combustion research that considers radiation employs an optically thin approximation despite its limited applicability to combustion processes. Assuming optical thinness can predict the radiative behavior of small laboratory flames reasonably well, but in larger flames or industrial systems tends to overpredict radiative heat losses by 50\% or more and underpredicts flame temperatures by over 100$^{\circ}$C \citep{Modest_2016}.

The specific correlations for the Planck Mean absorption coefficients generated by RadLib come from the Turbulent Nonpremixed Flame (TNF) Workshop, an ongoing collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. These particular correlations were chosen for RadLib's PM model because of the popularity and widespread use of the TNF Workshop radiation model in combustion research despite its limitations. Its temperature-dependent correlations are based on the RADCAL database \citep{Grosshandler_1993}, which has been largely superseded by more comprehensive and accurate databases, particularly HITRAN and HITEMP \citep{Rothman_2010}. The absorption coefficient correlations use fourth or fifth order polynomial fits in the range 300-2500 K. TNF Workshop literature shows that in cases where the optically thin approximation accurately describes the system, such as in simple hydrogen jet flames, the PM model produces accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, however, including most practical combustion scenarios of interest to scientists and engineers, the PM model does not produce accurate results, overpredicting radiative heat loss and underpredicting flame temperature by 100$^{\circ}$ or more \cite{Frank_2000,Zhu_2002,Coelho_2002}.

Because of these limitations, the authors strongly recommend that RadLib users carefully consider the implications of the PM model before using it in combustion simulations. It is implemented in RadLib primarily for comparison and evaluation against other radiation property models because of its popularity and widespread use. The examples included with the RadLib code and shown in Section \ref{s:Examples} demonstrate the deficiencies of the PM model in comparison to other radiation property models included with RadLib.

@@

The spectral line-weighted-sum-of-gray-gases (SLW) model represents a family of global approaches to radiative heat transfer in high-temperature gases that also includes absorption distribution functions (ADF) and full spectrum $k$-distribution (FSK) models, all of which are based on the same fundamental principle in modeling the gas absorption spectrum \cite{Solovjov_2016}. In order to extend their spectral models from uniform conditions (isothermal, homogeneous gases) to nonuniform conditions (non-isothermal, non-homogeneous gases), these models take a reference approach in which local gas states are corrected relative to a reference state. Reference approaches, however, generally lack consistent reference states and can yield significant errors in cases with large spatial temperature gradients \cite{Solovjov_2017}. The Rank Correlated SLW (RCSLW) model is a unique extension of the generalized SLW model that does not require a specified gas reference state.

The RCSLW model was chosen for RadLib for its accuracy and performance compared to other SLW and WSGG modeling approaches. A recent comparison of SLW modeling approaches revealed that ``the Rank Correlated SLW model is the most robust of all models, and demonstrates that it can achieve accurate solutions with as few as 3–5 gray gases" \citep{Badger_2019}. The literature indicates that the RCSLW model has not yet been successfully applied to turbulent combustion simulations \cite{Modest_2016,Coelho_2017}. 
Other SLW models have been applied to gas--soot mixtures \cite{Solovjov_2001,Solovjov_2001b} and high-temperature non-isothermal gases \cite{Solovjov_2011,Solovjov_2020} with results that suggest their applicability to combustion systems.

RadLib implements the RCSLW method presented by Solovjov et al. \cite{Solovjov_2017} as method I.2.2, which gives the best and most consistent results of the SLW methods discussed in the aforementioned study. A summary of the general SLW method and the RCSLW method are presented here, but additional information about these and various other SLW methods can be found in the literature \cite{Solovjov_2000, Solovjov_2001, Solovjov_2008, Solovjov_2011, Solovjov_2014, Solovjov_2016, Solovjov_2017, Webb_2018}.

>>>>

%% This list environment is used for the references in the
%% Program Summary
%%
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{float}
\usepackage[version=4]{mhchem}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{seqsplit}
\usepackage{enumitem}

@@

\newcommand{\BS}{\boldsymbol}
\setlist[itemize]{noitemsep}%, topsep=0pt}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
    RadLib is a C++ library of radiation property models that can be applied to variety of systems involving radiative heat transfer, including CFD simulations. RadLib includes three major radiation property models---Planck Mean (PM) absorption coefficients, the weighted sum of gray gases (WSGG) model, and the rank-correlation spectral line weighted-sum-of-gray-gases (RCSLW) model. RadLib includes C++, Python, and Fortran interfaces and can be expanded to include additional models.Several example cases illustrate the use of the models with an included ray-tracing solver and compare them in terms of accuracy relative to line-by-line solutions. A CFD example is provided of an ethylene burner in a compartment using Fire Dynamics Simulator (FDS). The computational cost of the models is compared. RadLib provides researchers with convenient access to validated radiation property models and a framework for further development. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%k%

{\bf PROGRAM SUMMARY}

@@

Modeling radiative heat transfer in participating media in high temperature systems is a challenging problem. Combustion is one of the key radiative tranfer applications, and combustion CFD is uniquely challenging because it involves numerous tightly coupled physical phenomena---including multicomponent mass transfer, convective and radiative heat transfer, turbulent fluid dynamics, multi-phase and particle flows, and complex chemical reaction kinetics---that span many orders of magnitude in both length and time scales. Combustion simulations are often complicated by a lack of accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost. Frequently, radiative heat transfer modeling in combustion systems is either neglected or oversimplified, resulting in significant errors flame temperatures \citep{Modest_2016}. Additionally, some late-stage flame phenomena, particularly soot behavior and flamesheet breakthrough, increasingly depend on and influence the magnitude of radiative heat losses and are difficult to simulate accurately without robust radiation modeling \cite{Modest_2013}. These problems motivated development of the RadLib library, which implements three radiation property models, with C++, Python, and Fortran interfaces, and provides a framework for additional models and further development, giving researchers convenient access to advanced radiation property modeling tools.

The core mechanisms of radiative phenomena are physically and mathematically complex. Unlike other modes of heat transfer, radiation is governed by a set of integro-differential equations that depend on both direction and wavelength, known as the radiative transfer equations (RTE). Computational radiation solvers require two distinct but interconnected parts: (1) a solution approach for the radiative transfer equations (RTE) and (2) a radiative property model. 
Complexity, implementation, and computational costs of these two parts can vary widely, depending on the models chosen. The optically thin model for solution of the RTE is trivial, while the discrete ordinates model (DOM) and Reverse Monte Carlo Ray Tracing (RMCRC) model, among others, are significantly more involved. Similarly for radiation property models, which range from simple correlations for mean absorption coefficients, to gray gas models such as the weighted sum of gray gas model (WSGG), to the more advanced Spectral Line WSGG (SLW) family of model, to full line-by-line (LBL) treatment. In practical engineering simulations, the solution of the RTE is often the most complex and costly part of radiative simulation, but advanced property models, such as the Rank Correlated SLW (RCSLW), discussed here, are nontrivial models to understand, implement, and validate.
RadLib addresses this second radiative property model part of radiation simulation by providing a set of implemented and validated radiation property models that can be applied to various solution approaches for the RTE.

For an absorbing, emitting, and non-scattering medium (considered here), the RTE for intensity $I$ at a specified wavenumber $\eta$ in unit direction $\mathbf{s}$ is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, $I_{b,\eta}$ is the spectral blackbody radiative intensity (or Planck function), $\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a path length along direction $\mathbf{s}$. Solving the RTE requires information about the radiative properties of the materials and media in the system. Radiative property data typically takes the form of radiation absorption coefficients $\kappa$, which are provided by a radiation property model. Gas absorption spectra consist of millions of spectral lines corresponding to the individual quantum mechanical energy levels of gas molecules, and depends on the local thermochemical state of the system (e.g., temperature, pressure, and composition). Radiation property models, including those implemented in RadLib, are typically developed using high-resolution spectral databases like HITRAN and HITEMP \cite{Rothman_2010} in order to calculate spectral absorption coefficients, thus reducing or simplifying the huge amount of spectral data by using correlations and modeling assumptions \citep{Zhang_2002b}.

The direct solution of the spectral RTE using so-called line-by-line calculations in which the RTE is solved for each individual spectral line is computationally prohibitive for most practical configurations and is used primarily for model validation. Some band models rely on estimated or averaged "line-of-sight" property values that cannot be directly incorporated into the RTE.
%VBS: citation here?
Modern band models require high-resolution databases and reorder emission levels rather than averaging across sequential absorption bands, but these cant lose accuracy when applied to nonhomogeneous media and may still require large numbers of RTE evaluations \citep{Modest_2016}.

Global models are an important class of radiation property models that use spectrally-integrated radiation properties \cite{Hottel_1967,Modest_2013,Modest_2016}. Assuming spectrally uniform (gray) properties over a wavenumber region $j$, we can integrate Eq.~\ref{e:RTE_general} over the fixed wavenumber ranges $\eta_j$ to obtain 
%
\begin{linenomath}
\begin{equation} \label{e:RTE_wsgg}
	\frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
\end{linenomath}
%
where $I_b$ is the blackbody intensity, $I_j$ is the radiative intensity in the region defined by gray gas $j$, $\kappa_j$ is the spectrally integrated absorption coefficient for gray gas $j$, and $a_j$ is a weighting factor defined by
%
\begin{linenomath}
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
\end{linenomath}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~\ref{e:RTE_wsgg} can be evaluated for each of the $n$ gray gases. Normally, an additional ``clear gas'', $j=0$, is included that represents transparent windows in the spectrum for which $\kappa_0=0$. This gray gas approach allows calculation of the total radiative intensity $I$ by summing the individual gray gas: 
%
\begin{linenomath}
\begin{equation}
	I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
\end{linenomath}
%

The heat flux $\mathbf{q}$ (in units of kW/m$^2$) and volumetric radiative heat source $Q$ (kW/m$^3$) can be calculated by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere and finding the divergence of the resulting vector field,
%
\begin{linenomath}
\begin{equation}
    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
\end{equation}
\end{linenomath}
%
\begin{linenomath}
\begin{equation}
    Q = -\nabla\cdot\mathbf{q}.
\end{equation}
\end{linenomath}
%
Global models are especially attractive in computational heat transfer applications because they can provide good accuracy while limiting computational costs.

With RadLib, we aim to provide researchers with a convenient access point for radiative property models that can be interfaced with various RTE solution approaches suitable for CFD applications. RadLib is a C++ library with optional Python and Fortran interfaces. RadLib calculates radiation absorption coefficients and weighting factors for use in global radiation models. Expansion to other radiative properties and model types is possible. Three models are fully implemented and validated. 
%All models use a common interface to facilitate researchers' access to and use of radiation modeling tools suitable for combustion CFD applications.

@@

At present, RadLib includes three radiation property models of varying complexity: Planck Mean absorption coefficients using the optically thin approximation, a weighted sum of gray gases (WSGG) model, and the rank-correlated spectral line weighted-sum-of-gray-gases (RCSLW) model. These models were chosen for their relavance to practical radiative heat transfer simulations. They represent increasing levels of complexity, computational cost, and accuracy. Each model is discussed in more detail in the following sections.

%Absorption coefficients and other radiative property data may take on different forms depending on the RTE solution approach, but can generally be classified as either spectral (wavenumber-dependent) or global (integrated over wavenumber). Spectral absorption coefficients can be applied to the spectral RTE (Eq.~\ref{e:RTE_general}), while global absorption coefficients are used in the spectrally-integrated RTE (Eq.~\ref{e:RTE_wsgg}). The type of absorption coefficient generated by each model and potential RTE integration strategies will also be discussed in the following sections. While solving the RTE is not the focus of this software, RadLib's example cases do employ a simple ray tracing solver to calculate the radiative heat flux and volumetric heat source profiles between two parallel planes. Refer to the example cases in Section \ref{s:Examples} for details.

Radiation absorption coefficients are typically calculated via correlations developed from high-resolution spectral databases like RADCAL, HITRAN, and HITEMP \citep{Grosshandler_1993,Rothman_2010} that relate spectral line strength to properties such as temperature, pressure, and gas composition \citep{Zhang_2002b}. At present, RadLib considers up to four gas species (H$_2$O, CO, CO$_2$, and CH$_4$), depending on the model, and soot. The property models implemented in RadLib apply only to non-scattering media.

@@

The Planck Mean (PM) model is the simplest of the three models implemented. This model assumes a single gray gas and computes the absorption coefficient as a function of temperature, pressure, and gas composition.

The particular version of the model implemented is from the Turbulent Nonpremixed Flames (TNF) workshop \cite{TNF}. This is a long-running collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. The workshop includes a number of laboratory jet flames that have been extensively researched for flame temperature, velocity, and composition profiles, and these serve as target flames for computational scientists to validate simulations and models. The PM model implemented here is the one advocated (and hence widely used) by the TNF workshop.

For each species, the PM model provides the absorption coefficient as fourth or fifth order polynomial fits in one or two temperature ranges to $\kappa(T)$ values computed from the RADCAL database \cite{Grosshandler_1993}. See \cite{TNF} for details and references. The PM model here considers the species \ce{CO2}, \ce{H2O}, \ce{CO}, \ce{CH4}. The last two species have lower absorption than the first two, but are included because \ce{CH4} and syngas (consisting of \ce{CO} and \ce{H2}) are common fuels in combustion applications and so are present in the TNF flames in high concentrations in the fuel stream. (\ce{CO} is also a common species in rich flame regions and is also included in the RCSLW model.)

The PM model is not as accurate as the other two models considered, and this is illustrated in Section~\ref{s:discussion}. 
It is implemented in RadLib primarily for comparison and evaluation against other radiation property models because of its relatively common use in some simulations.
The model has been reported to overpredict radiative losses and underpredict flame temperature by as much as 100 K or more \cite{Frank_2000,Zhu_2002,Coelho_2002}, though this will depend strongly on the geometric configuration. Accurate results have been reported for hydrogen jet flames \cite{Barlow_1999}.

The PM model is often used in conjunction with the optically-thin approximation for solving the RTE, in which case the radiative source term of the energy transport equation is particularly simple, and the overall computational cost of the radiative treatment is negligible in comparison to other aspects of a reacting flow simulation. Many smaller combustion applications are radiatively thin, and have smaller radiative heat loss. It is not uncommon for non-sooting laboratory-scale research jet flames to have radiative losses of around 5\%. At these levels, the accuracy of the radiative property model is not as crucial, and simpler models like the PM model are appropriate. 
% todo: show the RTE equation/source term

%Its temperature-dependent correlations are based on the RADCAL database \citep{Grosshandler_1993}, which has been largely superseded by more comprehensive and accurate databases, particularly HITRAN and HITEMP \citep{Rothman_2010}. The absorption coefficient correlations use fourth or fifth order polynomial fits in the range 300-2500 K. TNF Workshop literature shows that in cases where the optically thin approximation accurately describes the system, such as in simple hydrogen jet flames, the PM model produces accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, however, including most practical combustion scenarios of interest to scientists and engineers, the PM model does not produce accurate results, overpredicting radiative heat loss and underpredicting flame temperature by 100$^{\circ}$ or more \cite{Frank_2000,Zhu_2002,Coelho_2002}.

%RadLib's simplest radiation property model generates Planck Mean absorption coefficients under the assumption of optically thin media, from here on referred to as the Planck Mean (PM) model.
%The optically thin approximation assumes that there is little to no reabsorption of radiative energy in the system, or, in other words, that the system only emits radiation and does not absorb it. In combustion systems, optical thinness is associated with various other flame properties such as low radiative heat transfer relative to chemical heat release or convective heat transfer, negligible amounts of soot or other large molecules, and relatively small flame lengths. These flame attributes are not universally indicative of optical thinness, but are often used to justify neglecting or simplifying radiative heat transfer modeling. Assuming optical thinness, however, has the distinct advantage of simplifying the radiative source term (denoted $S_{rad}$) of the energy conservation equation to 
%\begin{equation}
%	S_{rad}=-4\kappa \sigma T^4,
%\end{equation}
%where $\kappa$ is the radiation absorption coefficient, $\sigma$ is the Stefan-Boltzmann constant, and $T$ is temperature. Using an optically thin approximation means that no RTE solution is required when calculating radiative source terms.
%
%In combustion systems, this simplification is convenient because it makes radiation modeling easy to implement and computationally inexpensive. The vast majority of computational combustion research that considers radiation employs an optically thin approximation despite its limited applicability to combustion processes. Assuming optical thinness can predict the radiative behavior of small laboratory flames reasonably well, but in larger flames or industrial systems tends to overpredict radiative heat losses by 50\% or more and underpredicts flame temperatures by over 100$^{\circ}$C \citep{Modest_2016}. 
%
%The specific correlations for the Planck Mean absorption coefficients generated by RadLib come from the Turbulent Nonpremixed Flame (TNF) Workshop, an ongoing collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. These particular correlations were chosen for RadLib's PM model because of the popularity and widespread use of the TNF Workshop radiation model in combustion research despite its limitations. Its temperature-dependent correlations are based on the RADCAL database \citep{Grosshandler_1993}, which has been largely superseded by more comprehensive and accurate databases, particularly HITRAN and HITEMP \citep{Rothman_2010}. The absorption coefficient correlations use fourth or fifth order polynomial fits in the range 300-2500 K. TNF Workshop literature shows that in cases where the optically thin approximation accurately describes the system, such as in simple hydrogen jet flames, the PM model produces accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, however, including most practical combustion scenarios of interest to scientists and engineers, the PM model does not produce accurate results, overpredicting radiative heat loss and underpredicting flame temperature by 100$^{\circ}$ or more \cite{Frank_2000,Zhu_2002,Coelho_2002}. 
%
%Because of these limitations, the authors strongly recommend that RadLib users carefully consider the implications of the PM model before using it in combustion simulations. 
%It is implemented in RadLib primarily for comparison and evaluation against other radiation property models because of its popularity and widespread use. 
%The examples included with the RadLib code and shown in Section \ref{s:Examples} demonstrate the deficiencies of the PM model in comparison to other radiation property models included with RadLib.

@@

The spectral line-weighted-sum-of-gray-gases (SLW) model represents a family of global approaches to radiative heat transfer in high-temperature gases that also includes the absorption distribution function (ADF) and full spectrum $k$-distribution (FSK) models, all of which are based on the same fundamental principle in modeling the gas absorption spectrum \cite{Solovjov_2016}. In order to extend their spectral models from isothermal, homogeneous composition gases to non-isothermal, non-homogeneous composition gases, these models use a reference approach in which radiative properties at local gas states are computed with respect to properties a at reference state. Such approaches, however, generally lack consistent reference states and can yield significant errors in cases with large spatial temperature gradients \cite{Solovjov_2017}. The Rank Correlated SLW (RCSLW) model is a unique extension of the generalized SLW model that does not require a specified gas reference state.

The RCSLW model was chosen for RadLib for its accuracy and performance compared to other SLW and WSGG modeling approaches. A recent comparison of SLW modeling approaches revealed that ``the Rank Correlated SLW model is the most robust of all models, and demonstrates that it can achieve accurate solutions with as few as 3–5 gray gases" \citep{Badger_2019}. The literature indicates that the RCSLW model has not yet been applied to turbulent combustion simulations \cite{Modest_2016,Coelho_2017}. We provide an example below using the Fire Dynamics Simulator (FDS) \cite{FDS}. 
Other SLW models have been applied to gas--soot mixtures \cite{Solovjov_2001,Solovjov_2001b} and high-temperature non-isothermal gases \cite{Solovjov_2011,Solovjov_2020} with results that suggest their good applicability to combustion systems.
